A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-11-03 latex/main.tex
Updates from Overleaf

To validate the accuracy of times for non-sample internal nodes inserted 
into the ARG
(inferred using tsdate), we compared the dates of first emergence of
Nextstrain clades against the Nextstrain SARS-CoV-2 tree (downloaded on 
2023-01-21). Figure \ref{fig:pango_vs_nextstrain_node_times} shows relatively
close agreement for most clades, with only three 
disagreeing by more than 28 days: 20C (an early lineage from 2020), 20E (Pango B.1.177 also known as EU1, emerging mid-2020), and 21B (Pango B.1.617.1 also known as Kappa, emerging in late 2020 to early 2021).

Then, to gauge internal consistency of the time estimation procedure, we
selected a subset of sample nodes, matching the distribution of node times 
and average number of children of the non-sample internal nodes. These
selected sample nodes were then re-dated, and the inferred dates compared 
against their recorded collection dates. The results in Figure 
\ref{fig:node_time_testing} demonstrate that although the inferred and true
times are generally very close, there is a clear bias with re-estimated times
being consistently older (by an average of 37 days). This is in line with 
the findings of other recent
work \cite{sanderson2021chronumental}, and is likely to be a consequence
of edges carrying more mutations than would be expected under a neutral
molecular clock prior (either due to errors in tree-building or genuinely varying mutation rates), as well as inaccuracies in the recording of 
sample dates.

>>>>

To validate the accuracy of times for non-sample internal nodes
inserted into the ARG (inferred using tsdate),
we compared the dates of first emergence of
Nextstrain clades against the Nextstrain SARS-CoV-2 tree
(downloaded on 2023-01-21).
Figure~\ref{fig:pango_vs_nextstrain_node_times} shows
relatively close agreement for most clades,
with only three disagreeing by more than 28 days:
20C (an early lineage from 2020),
20E (Pango B.1.177 also known as EU1, emerging mid-2020), and
21B (Pango B.1.617.1 also known as Kappa, emerging in late 2020 to early 2021).

Then, to gauge internal consistency of the time estimation procedure,
we selected a subset of sample nodes,
matching the distribution of node times  and
average number of children of the non-sample internal nodes.
These selected sample nodes were then re-dated, and
the inferred dates compared against their recorded collection dates.
The results in Figure~\ref{fig:node_time_testing} demonstrate that
although the inferred and true times are generally very close,
there is a clear bias with re-estimated times
being consistently older (by an average of 37 days).
This is in line with the findings of
other recent work \cite{sanderson2021chronumental}, and
is likely to be a consequence of edges
carrying more mutations than would be expected
under a neutral molecular clock prior
(either due to errors in tree-building or genuinely varying mutation rates),
as well as inaccuracies in the recording of sample dates.
